---
title: "TidyTuesday - Tour de France"
author: "Shemaiah S. Sotrah"
date: "4/8/2020"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(gganimate)
library(gifski)
library(transformr)
```

## Load Data

```{r, message=FALSE}
tdf_winners <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-04-07/tdf_winners.csv')

stage_data <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-04-07/stage_data.csv')

tdf_stages <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-04-07/tdf_stages.csv')

head(tdf_winners)
head(stage_data)
head(tdf_stages)
```

## Manipulate Data

```{r, message=FALSE}
 stage_data %>%
  filter(year %in% c("2002", "2003"), #just to compare the two most recent years
         rider %in% c("O'Grady Stuart", "Armstrong Lance")) %>% #compare winners
  separate(time, into = c("number", "label"), sep = "S") %>% #to avoid being identified as character
  separate(stage_results_id, into = c("label", "stage"), sep = "-") %>% #to avoid being identified as character
  select(- label) %>% #not informative column
  mutate(number = as.numeric(number)) %>% #changing both number and stage to numeric
  mutate(stage = as.numeric(stage)) -> winner
```

## Visualize Data

```{r, 3fig.width= 9, fig.height= 5}
ggplot(winner, aes(x = stage, y = number, fill = rider)) + 
  geom_col(position = "dodge") + #cols adjacent to each other instead of stacked
  facet_wrap(~year) + #compare two years side by side
  theme_classic() +
  scale_fill_manual(values = c("skyblue1", "sienna1")) +
  labs(x = "Race Stage",
       y = "Time (sec)",
       fill = "Rider",
       title = "Race Time Comparison for Tour de France",
       subtitle = "2002 and 2003") -> plot_winner

plot_winner + 
  transition_states(stage) + 
  shadow_mark()
```

```{r, fig.width= 9, fig.height= 5}
plot_winner
ggsave("plot_winner3.png", plot = plot_winner) #saved without animation
```

## Explanation

In 2002 and 2003, both Lance Armstrong and Stuart O'Grady competed in the Tour de France. However, in 2002, Armstrong won and O'Grady won the following year. For this reason, I thought it would be interesting the compare the individual riders' times in each stage of the race for the two consecutive years. Gganimate was a good option for this. It allowed me to isolate each stage to compare the stage times between the two riders. In this, way it made the information more digestible, especially when comparing over 2 years. 

