---
title: "GDPR Violations TidyTuesday"
author: "Shemaiah S. Sotrah"
date: "4/26/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(hrbrthemes)
```

## Load Data

```{r}
gdpr_violations <- readr::read_tsv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-04-21/gdpr_violations.tsv')
gdpr_text <- readr::read_tsv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-04-21/gdpr_text.tsv')

head(gdpr_violations)
head(gdpr_text)
```


## Manipulate Data

```{r}
gdpr_violations %>%
 filter(name == "Bulgaria") -> gdpr_bulgaria #only singles out Bulgaria

```

## Visualize Data

```{r}
ggplot(gdpr_bulgaria, aes(x = date, y = log(price))) + #log price so more comparable
  geom_segment(aes(x = date, y = log(price),
                   xend = date, #segments connect from date to point
                   yend = 0), #all start from price 0
                   color = "white") +
  geom_point(color = "green", #changes color of circle
             shape=20, #specific circle
             size = 3) + #increased size of circle
  theme_ft_rc() + #dark background
  labs(y = "log price in euros",
       title = "GDPR Violations in Bulgaria") +
  coord_flip() -> p1 #switches the x and y axis for more visible labels

p1

ggsave("gdpr_bulgaria.png", plot = p1)
```

## Explanation

  In this TidyTuesday, I used a lollipop chart. To me, it seemed similar to geom_col(), but it allowed me to have more than one data point for a given x and y. For example, the date 9/13/2019 had multiple prices, and this could neatly be shown with the conjunction of geom_point and geom_segment. Another feature seen in this plot was the flipping of the x and y axis. I thought this was incredibly useful in this case since labels in the x axis were longer. I also had fun with changing the theme to a darker background, which was something I had never tried before. It accented the data with brighter neon colors against the dark background. I did this using a new package that I discovers on github: hrbrthemes.  
